---
title: "Acquisition of spirantization across word boundaries in Ln Spanish by L1 speakers of German" 
date: November 14, 2024
date-format: long
author: Fernanda Barrientos
include-in-header: 
  text: |
    <style>
      .custom-small table {
        font-size: .75em
      }
    </style>
    <style>
      #title-slide .title {
        font-size: 2em;
        color: #b22222;
      }
      </style>
format:
  revealjs:
    slide-level: 2
    incremental: true
    theme: simple
    smaller: true
    footer: "Universität Konstanz | Spanish spirantization by L1 German speakers | Barrientos"
    slide-number: true
  pdf:
    pdf-engine: weasyprint
auto-stretch: false
fig-responsive: true
fig-asp: 0.6
fig-width: 7
fig-align: "center"
---

## Introduction

- Most studies regarding the acquisition of a second language’s (L2) phonology focus on segmental acquisition from a phonetic perspective (Gut 2009: 15), without engaging with formal phonological theory
- Exceptions relevant to this topic: Young-Scholten 1994, Zampini 1998, Lleó and Vogel 2003, Cabrelli Amaro 2017 
- Furthermore, very few studies in the acquisition of L2 Spanish by native (L1) German speakers have been carried out

## The Prosodic Hierarchy

- Nespor and Vogel (1986): phonological structure can be broken down into different domains
  - phonological utterance (U), 
  - intonational phrase (IP), 
  - phonological phrase (PP), 
  - phonological word (PW) 
  - foot (Ft) 
- Phonological rules apply within these domains, with variability across languages

## German 

- Lleó and Vogel (2004): German is a "demarcating" language
  - The focus is kept on the PW via demarcating strategies such as aspiration, devoicing, and glottal stop insertion
  - The latter, plus degemination, are strategies for syllable repair (Moosmüller 2015)
  - Standard German phonology does not resyllabify (Bergmann 2014), which blocks the possibility of applying phonological rules across word boundaries 
  - _Der Apfel_ is produced as [dɛɐ. ʔa.͡pfl] and not *[dɛ.ʀa.͡pfl] (i.e. the /r/ phoneme is implemented as a word-final, vocalized coda, and not as the onset of the following syllable

## Spanish

- Spanish, on the other hand, is a "grouping" language (Lleó and Vogel 2004): 
  - Phonological processes apply across longer constituents, where Spirantization is found within the IP
  - Spirantization is defined by the following rule:
    - [-cont, +voice] $\rightarrow$ [+cont] / [+cont] (##) __
    - ... Whereby /b d g/ are produced as fricatives (actually, approximants) after continuant sounds, with or without a word boundary in between; resyllabification occurs as a syllable repair strategy (Colina 2004, Hualde 2005)
    - _Abogado_ ('lawyer'): [aβoɣaðo]
    - But also: _una gata_ ('a female cat'): [una ɣata]
      
## Hypothesis

- Native speakers of German will transfer the demarcating pattern to Spanish
  - This means that they may apply the rule word-medially, but not word-initially
  - The degree of spirantization (from plosive to fricative to approximant) increases along with proficiency
    - This can be measured through intensity differences: if the /b d g/ sounds are produced with less constriction, they will be produced with more intensity

## Method

::: {.panel-tabset}

### Participants

- 26 late-learners of L1 German (mean age= 26; m=9; f=17) at different stages of acquisition
- Most of them reported high proficiency in English
- Their proficiency levels spanned from A1 to C1 (self-reported); A1 to B2 by placement test with CLOZE technique


### Materials

- Introduction prompts (where were you born, what's your age, what do you do, why do you want to learn Spanish)
- Three sets of images: food, activities, animals
- Two short texts: one narrative, one academic

### Procedure

- Placement test
- Sequence:
  - First task: introducing themselves
  - Second task: picture descriptions
  - Third task: read words embedded in carrier sentences
  - Fourth task: read two texts aloud
- Short breaks between tasks
- Participants had the chance to learn the words that they would need in advance, also those in the reading tasks

:::

## Data analysis

- The recordings were phonologically annotated with the forced-aligner FASE (Wilbanks, 2022). 
- We analyzed the instances of phonological /bdg/ following Cabrelli Amaro (2017) by measuring the differences in intensity (dB) between the consonant and its following vowel: the lower the difference, the higher the level of spirantization
- Words with orthographic <v> and with /bgd/ following obstruents or pauses were excluded
- We used the percentage of correct responses in the placement test as a continuous measure for proficiency

## Results
::: {.nonincremental}

::: {style="font-size: 0.8em"}
- Only the last task has been analyzed so far...

```{r eval=TRUE, echo=FALSE, output=FALSE}
library(lme4)
library(car)
library(emmeans)
newsounds <- read.csv("/home/fernanda/Desktop/Paths/newsounds.txt", stringsAsFactors=TRUE)
#Run regs
newsounds2<-subset(newsounds, Context!= "final")
aver<-lmer(IntensityDifference ~ Consonant * Context * test_perc+ (1|Subj) + (1|Word), data=newsounds2)
Anova(aver, type="III")
```

- We tested the effect of context (word-initial, word-medial), proficiency (as percentage), and consonant (/b d g/) on intensity differences
- Three-way interaction between all predictors ($\chi^2$ = 6.4719, df= 2, p<0.05)

:::

```{r eval=TRUE, echo=FALSE, output=TRUE}
library(ggplot2)
pcons <- ggplot(newsounds2, aes(x = Consonant, y = IntensityDifference, fill = Context)) +  
  geom_boxplot(outlier.size=2) +
  theme_minimal() 
print(pcons)
```

:::

## Results

::: {.nonincremental}

::: {style="font-size: 0.8em"}

```{r eval=TRUE, echo=FALSE, output=FALSE}
conts<-emmeans(aver, ~ test_perc * Context * Consonant)
pairs(conts, simple="Context")
```

- /g/ shows the least spirantization across contexts and levels of proficiency
- /b-d/ show more word-medial spirantization as proficiency increases.
- Word-initially, the degree of spirantization for /b-d/ increases with proficiency, but without reaching the values seen word-medially. 

:::

```{r eval=TRUE, echo=FALSE, output=TRUE}
library(emmeans)
emmip(aver, Consonant ~ test_perc | Context, cov.reduce=range, xlab="% Correct answers in test", ylab= "Intensity Different (dB)", dotarg = list(color = "black"))
```

:::

## Discussion

- We conclude that L1 German speakers tend towards acquiring a target-like phonology in Spanish as proficiency increases
  - First, spirantization is seen mostly in word-medial contexts
  - With experience, the domain of application is extended to word-initial contexts
- /g/ offers more difficulty (why?)
- a native phonology without resyllabification may block the application of postlexical rules across word boundaries even in experienced learners
- Is this German or just universal L2 behaviour?

## References

::: {style="font-size: 0.7em"}

Bergmann, P. (2014). Reduction and deletion of glottal stops and geminates at phonological word boundaries in German compounds: Effects of word frequency and accentuation. In J. C. Reina & R. Szczepaniak (Eds.), Syllable and Word Languages (pp. 251–278). De Gruyter. https://doi.org/10.1515/9783110346992.251

Cabrelli Amaro, J. (2017). The role of prosodic structure in the L2 acquisition of Spanish stop lenition. Second Language Research, 33(2), 233–269. https://doi.org/10.1177/0267658316687356

Colina, S. (2006). Optimality-theoretic advances in our understanding of Spanish syllable structure. In F. Martínez-Gil & S. Colina (Eds.), Optimality-Theoretic studies in Spanish Phonology (pp. 172–204). John Benjamins.

Hualde, J. I. (2005). The sounds of Spanish. CUP.

Moosmüller, S. (2015). The interaction of prosody and phonotactics: Resyllabification in three varieties of German. Italian Journal of Linguistics, 27(1), 111–132.

Lleó, C., & Vogel, I. (2004). Learning new segments and reducing domains in German L2 phonology: The role of the Prosodic Hierarchy. International Journal of Bilingualism, 8(1), 79-102.

Wilbanks, E. (2022). faseAlign (Version 1.1.14) [Computer software]. https://github.com/EricWilbanks/faseAlign

Young-Scholten, M. (1994). On positive evidence and ultimate attainment in L2 phonology. Second Language Research, 10(3), 193–214.

Zampini, M. (1998). "L2 Spanish spirantization: A prosodic analysis and pedagogical implications." Hispanic Linguistics 10(1), 154-188.

:::

##

::: {style="font-size: 3em; text-align: center"}

Thank you!

:::

::: {style= "text-align: center"}

fernanda.barrientos-contreras@uni-konstanz.de